Hybrid platforms and bargaining power

### 2023-12-06 src/paper/application.tex
Add a useful lemma

\paragraph{One-sided case} In the first setting, the platform and the fringe firms bargain over the total profits of the platform and the fringe taken together, and the consumers are not included in the negotiations.
The following assumption describes this case.
\begin{assumption}
    \label{ass:profit_sharing}
    Consider the following cooperative game $\mathcal{G} = (\mathcal{N}, v)$.
    The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i\}, i \in [0, N_F]$.
    The value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
    \begin{align*}
        v(S) = \begin{cases}
            0 & \text{if } P \notin S \\
            N_C f(N_P, n_F(S)) & \text{if } P \in S
        \end{cases},
    \end{align*}
    where $n_F(S)$ is the measure of fringe firms in $S$.
    Each type of player gets their Shapley value in this game:
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= \varphi_P(\mathcal{G}), \\
        \pi_F(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}).
    \end{align*}
\end{assumption}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \begin{tikzpicture}[scale=0.8]
            \centering
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1, samples at={0, 0.02, ..., 0.98, 1},
                    xtick={0, 1}, ytick=\empty, xlabel={$s$}]
                \addplot[name path=f, black, thick] {x^2};
                \node[anchor= east] at (axis cs: .9, .9^2) {$N_C f(N_P, sN_F)$};
                \path[name path=bottom] (axis cs:0,0) -- (axis cs:1,0);
                \path[name path=top] (axis cs:0,1) -- (axis cs:1,1);

@@

                \node[anchor=north west] at (axis cs: .05, .95) {Fringe ($F$)};
                \node[anchor=south east] at (axis cs: .95, .05) {Platform ($P$)};
            \end{axis}
        \end{tikzpicture}
        \caption{Fringe firms are complements}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1, samples at={0, 0.02, ..., 0.98, 1},
                xtick={0, 1}, ytick=\empty, xlabel={$s$}]
                \addplot[name path=f, black, thick] {x^0.5};
                \node[anchor=north west] at (axis cs: .6, .6^0.5) {$N_C \tilde{f}(N_P, sN_F)$};
                \path[name path=bottom] (axis cs:0,0) -- (axis cs:1,0);
                \path[name path=top] (axis cs:0,1) -- (axis cs:1,1);

@@

\paragraph{Two-sided case} Alternatively, in certain settings, it might be appropriate to assume that the consumers (or, more generally, the entities on the other side of the market) also engage in the bargaining process.
The cooperative game describing this situation is formalized in \cref{ass:profit_sharing_two_sided}.
\begin{assumption}
    \label{ass:profit_sharing_two_sided}
    Consider the following cooperative game $\mathcal{G} = (\mathcal{N}, v)$.
    The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i, C_j\}, i \in [0, N_F], j \in [0, N_C]$.
    The value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
    \begin{align*}
        v(S) = \begin{cases}
            0 & \text{if } P \notin S \\
            n_C(S) f(N_P, n_F(S)) & \text{if } P \in S
        \end{cases},
    \end{align*}
    where $n_F(S)$ and $n_C(S)$ is the measure of fringe firms and consumers, respectively, in $S$.
    Each type of player gets their Shapley value in this game:
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= \varphi_P(\mathcal{G}), \\
        \pi_F(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}), \\
        \pi_C(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}).
    \end{align*}
\end{assumption}

@@

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[xscale=2,yscale=2,samples=200]
        \draw[->] (0,0) -- (0,2);
        \draw[->] (0,0) -- (5.5,0) node[below] {$n_F$};

        \draw[name path=piF,domain=0.01:5.5,variable=\x,black,thick] plot ({\x},{8*((\x)/(0.3+\x))-8*((\x-0.3*ln(\x+0.3)-0.361)/(\x))});
        \draw[name path=IF,domain=0.01:5.5,variable=\x,black,dashed,thick] plot ({\x},{(\x)/3});
        \node[anchor=north east] at (5,1) {$\pi_F(N_P, N_F)$};
        \node[anchor=south east] at (5,5/3) {$I_F N_F$};

        \fill [name intersections={of=piF and IF, name=i, total=\t}]
            [black] (i-2) circle (1pt);
        \draw[dashed] (i-2) -- (i-2 |- 0,0) node[below] {$N_F^*$};
    \end{tikzpicture}
    \caption{An example free entry equilibrium under \cref{ass:single_crossing}}
    \label{fig:equilibrium}
\end{figure}

@@

                \node[anchor=north west] at (axis cs: .05, 1.17) {Fringe ($F$)};
                \node[anchor=south east] at (axis cs: .95, .05) {Platform ($P$)};
            \end{axis}
        \end{tikzpicture}
        \caption{Profit shares with $N_P = 0.5$}
    \end{subfigure}
    \caption{Illustration of \cref{prop:share_of_fringe} (fringe firms are substitutes, $f(N_P, N_F) = \sqrt{N_P + N_F}$)}
    \label{fig:increase_N_P_fringe}
\end{figure}

Now let us turn to equilibrium entry as a function of the platform's product variety.
The previous proposition implies an almost immediate corollary regarding the equilibrium number of fringe entrants.
\begin{corollary}
    \label{cor:fringe_entry}
    Assume that $f, w$ are twice continuously differentiable.
    Let $N_F^*$ denote the equilibrium number of fringe firms.
    Let us also assume that $N_F^* > 0$.
    Furthermore, let us assume that
    \begin{align*}
        \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} < 0 \quad \forall n_F \leq N_F \\
        \frac{\partial^2 f(n_P, n_F)}{\partial n_F^2} < 0 \quad \forall n_F \leq N_F.
    \end{align*}
    Then the equilibrium number of fringe firms is also differentiable and
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < 0.
    \end{align*}
\end{corollary}
That is, the equilibrium number of entrants increases as a response to an increase in $N_P$ if the fringe firms are mostly complements, and decreases if they are mostly substitutes.
The reason simply lies in the concave or hump-shaped fringe profit function.
If $N_F^*$ is an equilibrium, then fringe profits minus entry costs are strictly positive for all $N_F \leq N_F^*$ and strictly negative for all $N_F > N_F^*$.
Therefore, if an increase in $N_P$ decreases the fringe's profits for every $N_F \geq 0$, equilibrium can be restored by decreasing the number of fringe entrants, and vice versa for the other case (\cref{fig:comparative_N_F}).

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[xscale=2,yscale=2,samples=200]
        \draw[->] (0,0) -- (0,2);
        \draw[->] (0,0) -- (5.5,0) node[below] {$n_F$};

@@

        \node[anchor=south] at (5,1) {$\pi_F(N_P, n_F)$};
        \node[anchor=north] at (5,0.6) {$\pi_F(N_P', n_F)$};
        \node[anchor=south east] at (5,5/3) {$I_F n_F$};

        \fill [name intersections={of=piF and IF, name=i, total=\t}]
            [black] (i-2) circle (1pt);
        \draw[dashed] (i-2) -- (i-2 |- 0,0) node[below] {$N_F^*$};
        \fill [name intersections={of=piF2 and IF, name=i2, total=\t}]
            [black] (i2-2) circle (1pt);
        \draw[dashed] (i2-2) -- (i2-2 |- 0,0) node[below] {$N_F^{*'}$};

        \draw[->] (i-2 |- 0,0.5) -- (i2-2 |- 0,0.5);
    \end{tikzpicture}
    \caption{Illustration of comparative statics for equilibrium entry}
    \label{fig:comparative_N_F}
\end{figure}

Finally, let us conclude this section by establishing a stronger result for a more restrictive class of profit functions.
In the following, I assume that the profit function has a specific, additive form.
Intuitively, this means that -- in terms of total profits generated -- the platform's and the fringe firms' products are substitutable to each other according to some constant ratio.\footnote{
    Note that it does not mean that their products are substitutes in the usual, demand theory sense.
    They still might be complements to each other.
}
\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let $f$ have the following, additive form:
    \begin{align*}
        f(N_P, N_F) = g(\alpha N_P + \beta N_F)
    \end{align*}
    where $\alpha, \beta > 0$.
    Also assume that $g$ is twice differentiable and that $g'' < 0$.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
\end{proposition}
As opposed to the previous results, this proposition tells us something about the total profits generated by the platform and the fringe firms as a function of the platform's product variety, and under equilibrium entry behavior by the fringe firms.
In particular, it states that total profits increase if the total profit function is convex (products are substitutes), and decreases otherwise.
This result has particularly strong implications in models where total profits (or total product variety) has a monotone relationship with consumer welfare, such as those in \textcite{anderson2020aggregative}.
\Cref{sec:example} also presents a model in which this is the case.

@@

Let us now add more structure to the model, and microfound the profit function $\Pi$.
The model in this section is heavily inspired by the logit-like demand structure and hybrid platform setting in  \textcite{anderson2021hybrid}.
The setting is a (possibly hybrid) platform market with monopolistically competing fringe firms.

@@

@@

    Therefore, the initial assumption that $\frac{\partial N_F^*}{\partial N_P} \geq -\frac{\alpha}{\beta}$ was wrong.
    As we have from \cref{cor:fringe_entry} that $\frac{\partial N_F^*}{\partial N_P}$ exists, it must be the case that
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < \frac{\alpha}{\beta}
    \end{align*}
\end{proof}

>>>>

\paragraph{One-sided case} In the first setting, the platform and the fringe firms bargain over the total profits of the platform and the fringe taken together, and the consumers are not included in the negotiations.
The following assumption describes this case.
\begin{assumption}
    \label{ass:profit_sharing}
    Consider the following cooperative game $\mathcal{G} = (\mathcal{N}, v)$.
    The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i\}, i \in [0, N_F]$.
    The value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
    \begin{align*}
        v(S) = \begin{cases}
            0 & \text{if } P \notin S \\
            N_C f(N_P, n_F(S)) & \text{if } P \in S
        \end{cases},
    \end{align*}
    where $n_F(S)$ is the measure of fringe firms in $S$.
    Each type of player gets their Shapley value in this game:
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= \varphi_P(\mathcal{G}), \\
        \pi_F(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}),
    \end{align*}
    where $\varphi_i(\mathcal{G})$ denotes the Shapley value of player(s) $i$ in game $\mathcal{G}$.
\end{assumption}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \begin{tikzpicture}[scale=0.8]
            \centering
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1, samples at={0, 0.02, ..., 0.98, 1},
                    xtick={0, 1}, ytick=\empty, xlabel={$s$}]
                \addplot[name path=f, black, thick] {x^2};
                \node[anchor= east] at (axis cs: .5, .5^2) {$N_C f(N_P, sN_F)$};
                \path[name path=bottom] (axis cs:0,0) -- (axis cs:1,0);
                \path[name path=top] (axis cs:0,1) -- (axis cs:1,1);

@@

                \node[anchor=north west] at (axis cs: .05, .95) {Fringe ($F$)};
                \node[anchor=south east] at (axis cs: .95, .05) {Platform ($P$)};
            \end{axis}
        \end{tikzpicture}
        \caption{Fringe firms are complements}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1, samples at={0, 0.02, ..., 0.98, 1},
                xtick={0, 1}, ytick=\empty, xlabel={$s$}]
                \addplot[name path=f, black, thick] {x^0.5};
                \node[anchor=north west] at (axis cs: .5, .5^0.5) {$N_C \tilde{f}(N_P, sN_F)$};
                \path[name path=bottom] (axis cs:0,0) -- (axis cs:1,0);
                \path[name path=top] (axis cs:0,1) -- (axis cs:1,1);

@@

\paragraph{Two-sided case} Alternatively, in certain settings, it might be appropriate to assume that the consumers (or, more generally, the entities on the other side of the market) also engage in the bargaining process.
The cooperative game describing this situation is formalized in \cref{ass:profit_sharing_two_sided}.
\begin{assumption}
    \label{ass:profit_sharing_two_sided}
    Consider the following cooperative game $\mathcal{G} = (\mathcal{N}, v)$.
    The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i, C_j\}, i \in [0, N_F], j \in [0, N_C]$.
    The value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
    \begin{align*}
        v(S) = \begin{cases}
            0 & \text{if } P \notin S \\
            n_C(S) f(N_P, n_F(S)) & \text{if } P \in S
        \end{cases},
    \end{align*}
    where $n_F(S)$ and $n_C(S)$ is the measure of fringe firms and consumers, respectively, in $S$.
    Each type of player gets their Shapley value in this game:
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= \varphi_P(\mathcal{G}), \\
        \pi_F(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}), \\
        \pi_C(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}),
    \end{align*}
    where $\varphi_i(\mathcal{G})$ denotes the Shapley value of player(s) $i$ in game $\mathcal{G}$.
\end{assumption}

@@

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[xscale=2,yscale=2,samples=200]
        \draw[->] (0,0) -- (0,2);
        \draw[->] (0,0) -- (5.5,0) node[below] {$N_F$};

        \draw[name path=piF,domain=0.01:5.5,variable=\x,black,thick] plot ({\x},{8*((\x)/(0.3+\x))-8*((\x-0.3*ln(\x+0.3)-0.361)/(\x))});
        \draw[name path=IF,domain=0.01:5.5,variable=\x,black,dashed,thick] plot ({\x},{(\x)/3});
        \node[anchor=north east] at (5,1) {$\pi_F(N_P, N_F, N_C)$};
        \node[anchor=south east] at (5,5/3) {$I_F N_F$};

        \fill [name intersections={of=piF and IF, name=i, total=\t}]
            [black] (i-2) circle (1pt);
        \draw[dashed] (i-2) -- (i-2 |- 0,0) node[below] {$N_F^*$};
    \end{tikzpicture}
    \caption{An example free entry equilibrium. \Cref{ass:single_crossing} states that the fringe profit function is concave or hump-shaped in $N_F$, guaranteeing at most one intersection with the linear entry cost function.}
    \label{fig:equilibrium}
\end{figure}

@@

                \node[anchor=north west] at (axis cs: .05, 1.17) {Fringe ($F$)};
                \node[anchor=south east] at (axis cs: .95, .05) {Platform ($P$)};
            \end{axis}
        \end{tikzpicture}
        \caption{Profit shares with $N_P = 0.5$}
    \end{subfigure}
    \caption{Illustration of \cref{prop:share_of_fringe} when fringe and platforrm products are substitutes. An increase in the platform's product variety increases total profits, yet, the fringe's share decreases in absolute terms.}
    \label{fig:increase_N_P_fringe}
\end{figure}

Now let us turn to equilibrium entry as a function of the platform's product variety.
The previous proposition implies an almost immediate corollary regarding the equilibrium number of fringe entrants.
\begin{corollary}
    \label{cor:fringe_entry}
    Assume that $f, w$ are twice continuously differentiable.
    Let $N_F^*$ denote the equilibrium number of fringe firms.
    Let us also assume that $N_F^* > 0$.
    Furthermore, let us assume that
    \begin{align*}
        \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} < 0 \quad \forall n_F \leq N_F \\
        \frac{\partial^2 f(n_P, n_F)}{\partial n_F^2} < 0 \quad \forall n_F \leq N_F.
    \end{align*}
    Then the equilibrium number of fringe firms is also differentiable and
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < 0.
    \end{align*}
\end{corollary}
That is, the equilibrium number of entrants increases as a response to an increase in $N_P$ if the fringe firms are mostly complements, and decreases if they are mostly substitutes.
The underlyong reason is again the concave or hump-shaped fringe profit function.
If $N_F^*$ is an equilibrium, then fringe profits minus entry costs are strictly positive for all $N_F \leq N_F^*$ and strictly negative for all $N_F > N_F^*$.
Therefore, if an increase in $N_P$ decreases the fringe's profits for every $N_F \geq 0$, equilibrium can be restored by decreasing the number of fringe entrants, and vice versa for the other case (\cref{fig:comparative_N_F}).

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[xscale=2,yscale=2,samples=200]
        \draw[->] (0,0) -- (0,2);
        \draw[->] (0,0) -- (5.5,0) node[below] {$N_F$};

@@

        \node[anchor=south] at (5,1) {$\pi_F(N_P, N_F, N_C)$};
        \node[anchor=north] at (5,0.6) {$\pi_F(N_P', N_F, N_C)$};
        \node[anchor=south east] at (5,5/3) {$I_F n_F$};

        \fill [name intersections={of=piF and IF, name=i, total=\t}]
            [black] (i-2) circle (1pt);
        \draw[dashed] (i-2) -- (i-2 |- 0,0) node[below] {$N_F^*$};
        \fill [name intersections={of=piF2 and IF, name=i2, total=\t}]
            [black] (i2-2) circle (1pt);
        \draw[dashed] (i2-2) -- (i2-2 |- 0,0) node[below] {${N_F^*}'$};

        \draw[->] (i-2 |- 0,0.5) -- (i2-2 |- 0,0.5);
    \end{tikzpicture}
    \caption{Illustration of comparative statics for equilibrium entry. If the change in $N_P$ decreases the fringe's profits for every $N_F \geq 0$, equilibrium can be restored by decreasing the number of fringe entrants, and vice versa for the other case.}
    \label{fig:comparative_N_F}
\end{figure}

Finally, let us conclude this section by establishing a stronger result for a more restrictive class of profit functions.
In the following, I assume that the profit function has a specific, additive form.
Intuitively, this means that -- in terms of total profits generated -- the platform's and the fringe firms' products are substitutable to each other according to some constant ratio.\footnote{
    Note that it does not mean that their products are substitutes in the usual, demand theory sense.
    They still might be complements to each other.
}
\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let $f$ have the following, additive form:
    \begin{align*}
        f(N_P, N_F) = g(\alpha N_P + \beta N_F)
    \end{align*}
    where $\alpha, \beta > 0$.
    Also assume that $g$ is twice differentiable and that $g'' < 0$.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
    Furthermore,
    \begin{align*}
        \frac{\partial \Pi(N_P, N_F^*(N_P), N_C)}{\partial N_P} < 0.
    \end{align*}
\end{proposition}

The first part of this proposition is a stronger version of \cref{cor:fringe_entry}.
It states that, not only does the equilibrium number of fringe firms decrease as a response to an increase in $N_P$, but it establishes a lower bound for this decrease.
And, as the second part shows, this bound is sufficient to guarantee that the total size of the pie ($\Pi(N_P, N_F, N_C)$) also decreases in equilibrium.

This result has particularly strong implications in models where total profits (or total product variety) has a monotone relationship with consumer welfare, such as those in \textcite{anderson2020aggregative}.
\Cref{sec:example} also presents a demand system which satisfies this property.

@@

Let us now add more structure to the model, and microfound the profit function $\Pi$.
The model in this section is heavily inspired by the logit-like demand structure and hybrid platform setting in \textcite{anderson2021hybrid}.
The setting can be thought of as a (possibly hybrid) platform market with monopolistically competing fringe firms.

@@

\begin{lemma}
    \label{lem:single_crossing_affine_transformation}
    Let $g: \mathbb{R}^+ \to \mathbb{R}$ be a twice continuously differentiable function.
    Define
    \begin{align*}
        G(x) = \int_0^x g(sx) \ds.
    \end{align*}
    Assume that for any $x \geq 0$, $G'(x) < 0$ or $G''(x) < 0$.

    Let $h(x) = c g(a + bx)$ for some $a, c > 0$ and $b > 0$.
    Then the function $H(x) = \int_0^1 h(sx) \ds$ satisfies the same conditions: for any $x \geq 0$, $h'(x) < 0$ or $h''(x) < 0$.
\end{lemma}
\begin{proof}[Proof of \cref{lem:single_crossing_affine_transformation}]
    Differentiating $H$ yields
    \begin{align*}
        H'(x) &= \int_0^1 h'(sx) \ds \\
              &= \int_0^1 c b g'(a + bsx) \ds \\
              &= c b G'(a + bsx).
    \end{align*}
    Similarly, the second derivative is
    \begin{align*}
        H''(x) &= \int_0^1 h''(sx) \ds \\
               &= \int_0^1 c b^2 g''(a + bsx) \ds \\
               &= c b^2 G''(a + bsx).
    \end{align*}
    Now, for any $x \geq 0$, $y = a + bsx \geq 0$, therefore $G'(y) < 0$ or $G''(y) < 0$.
    As $c, b > 0$, this implies that $H'(x) < 0$ or $H''(x) < 0$.
\end{proof}

@@

    Therefore, the initial assumption that $\frac{\partial N_F^*}{\partial N_P} \geq -\frac{\alpha}{\beta}$ was wrong.
    As we have from \cref{cor:fringe_entry} that $\frac{\partial N_F^*}{\partial N_P}$ exists, it must be the case that
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}

    To prove the second part of the proposition, differentiate total profits with respect to $N_P$:
    \begin{align*}
        \frac{\partial \Pi(N_P, N_F^*(N_P), N_C)}{\partial N_F} &= \frac{\partial}{\partial N_F} N_C g(\alpha N_P + \beta N_F^*(N_P)) \\
        &= \underbrace{N_C g'(\alpha N_P + \beta N_F^*(N_P))}_{> 0} \left[ \alpha + \beta \frac{\partial N_F^*(N_P)}{\partial N_P} \right] \\
        &< N_C g'(\alpha N_P + \beta N_F^*(N_P)) \left[ \alpha + \beta \left( -\frac{\alpha}{\beta} \right) \right] \\
        &= 0.
    \end{align*}
\end{proof}
